\Large \bf FlexAlloc: Dynamic Memory Partitioning\\ between the Corevisor and Hostvisor in SeKVM

### 2023-12-24 paper.tex
Get paper to compile + Minor figure changes

% to be able to draw some self-contained figs
\usepackage{tikz}
\usepackage{amsmath}

@@

Although the metadata structs have a much smaller memory footprint than the stage-2 page tables,
we designed and partially implemented a fully generalized dynamic allocation interface that
could be used to support dynamic allocation for all metadata structs which we call the
Dynamic Page Manager (DPM). DPM handles all of the work required to receive and return
memory from the hostvisor allowing other "services" in EL2 to exchange memory with DPM using
a very similiar interface to the alloc_pages and free_pages interface available in EL1. We describe
this interface in detail in Section 4.

@@

\begin{figure}[h!tbp]
\caption{SeKVM VM Stage-2 Page Table Region Layout, Statically Allocated}
\label{fig:sekvmvmregions}
\centering
    \includegraphics{sekvm_vm_region}
\end{figure}

To get around this limitation, there are two possible approaches. The first approach
is to halve the size of the VM's stage-2 page table, so that it can fit within
a single contiguous 4M region of memory. Alternatively, we can fragment the VM's stage-2
page table such that it can be composed of multiple non-contiguous regions of memory, whose sizes in total add up to 8M. We
choose to implement the latter approach to avoid degrading VM performance, and to avoid
breaking the current implementation of the buddy allocator. In
particular, we fragment the VM's stage-2 page table so that it can be composed
of eight distinct 1M regions of memory. Each of these 1M regions must be contiguous,
but each region is distinct and need not be allocated contiguously with the
other 1M regions. As a result, the hostvisor can easily use \textit{alloc\_pages}
to allocate the memory needed for a guest VM's stage-2 page tables.

@@

\begin{figure}[h!tbp]
\caption{FlexAlloc: First 1MB Region Metadata Setup}
\label{fig:firstmbregion}
\centering
    \includegraphics{firstmbregion}
\end{figure}

\begin{figure}[h!tbp]
\caption{FlexAlloc: Second to Eigth 1MB Region Metadata Setup}
\label{fig:otherregions}
\centering
    \includegraphics{otherregions}
\end{figure}

Each level's iterators have a set number of regions from which to pull additional pages from. Lower
levels in the table have larger regions than higher levels in the table. When the current iterator
over a given region is exhausted, the iterator will jump to the next available region and pick up
where it left off.

@@

    // If we've used up all the pages and we
    // have another region, then use the new region
    if (used_pages == PTE_ITER_MAX_PAGES &&
        pool_index + 1 < PTE_USED_ITER_COUNT) {
        pool_index += 1;
        el2_data->vm_info[vmid].
          pte_pool_index = pool_index;

        // Set the pool_start to the next available
        // PTE iterator
        pool_start = el2_data->vm_info[vmid].
          pte_pool_starts[pool_index];
        used_pages = el2_data->vm_info[vmid].
          pte_used_pages_vm[pool_index];
    }

    // Make sure we have not exhausted all pages
    // across all regions
    if (used_pages == PTE_ITER_MAX_PAGES &&
        pool_index == PTE_USED_ITER_COUNT - 1)
        {
        __hyp_panic();
    }

    // Return the physical address of the next
    // page that this iterator can allocate
    return pool_start +
      (used_pages * PAGE_SIZE);
}
\end{verbatim}

>>>>

% to be able to draw some self-contained figs
\usepackage{tikz}
\usepackage{amsmath}
\usepackage{caption}

@@

Although the metadata structs have a much smaller memory footprint than the stage-2 page tables,
we designed and partially implemented a fully generalized dynamic allocation interface that
could be used to support dynamic allocation for all metadata structs which we call the
Dynamic Page Manager (DPM). DPM handles all of the work required to receive and return
memory from the hostvisor allowing other "services" in EL2 to exchange memory with DPM using
a very similiar interface to the alloc\_pages and free\_pages interface available in EL1. We describe
this interface in detail in Section 4.

@@

\begin{figure}[h!tbp]
\centering
\captionsetup{justification=centering}
\includegraphics{sekvm_vm_region}
\caption{SeKVM: Layout of Statically-Allocated\\Region for VM Stage-2 Page Tables}
\label{fig:sekvmvmregions}
\end{figure}

\bigskip To get around this limitation, there are two possible approaches. The first approach
is to halve the size of the VM's stage-2 page table, so that it can fit within
a single contiguous 4M region of memory. Alternatively, we can fragment the VM's stage-2
page table such that it can be composed of multiple non-contiguous regions of memory, whose sizes in total add up to 8M. We
choose to implement the latter approach to avoid degrading VM performance, and to avoid
breaking the current implementation of the buddy allocator. In
particular, we fragment the VM's stage-2 page table so that it can be composed
of eight distinct 1M regions of memory. Each of these 1M regions must be contiguous,
but each region is distinct and need not be allocated contiguously with the
other 1M regions. As a result, the hostvisor can easily use \textit{alloc\_pages}
to allocate the memory needed for a guest VM's stage-2 page tables.

@@

\begin{figure}[h!tbp]
\centering
\captionsetup{justification=centering}
\includegraphics{firstmbregion}
\caption{FlexAlloc: Metadata Setup\\for First 1MB Region}
\label{fig:firstmbregion}
\end{figure}

\begin{figure}[h!tbp]
\centering
\captionsetup{justification=centering}
\includegraphics{otherregions}
\caption{FlexAlloc: Metadata Setup\\for Second to Eigth 1MB Regions}
\label{fig:otherregions}
\end{figure}

\bigskip Each level's iterators have a set number of regions from which to pull additional pages from. Lower
levels in the table have larger regions than higher levels in the table. When the current iterator
over a given region is exhausted, the iterator will jump to the next available region and pick up
where it left off.

@@

    // If we've used up all the pages and we
    // have another region, then use the new
    // region.
    if (used_pages == PTE_ITER_MAX_PAGES &&
        pool_index + 1 <
        PTE_USED_ITER_COUNT) {
        pool_index += 1;
        el2_data->vm_info[vmid].
          pte_pool_index = pool_index;

        // Set the pool_start to the next
        // available PTE iterator.
        pool_start = el2_data->vm_info[vmid].
          pte_pool_starts[pool_index];
        used_pages = el2_data->vm_info[vmid].
          pte_used_pages_vm[pool_index];
    }

    // Make sure we have not exhausted all
    // pages across all regions.
    if (used_pages == PTE_ITER_MAX_PAGES &&
            pool_index ==
            PTE_USED_ITER_COUNT - 1) {
        __hyp_panic();
    }

    // Return the physical address of the
    // next page that this iterator can
    // allocate.
    return pool_start +
      (used_pages * PAGE_SIZE);
}
\end{verbatim}
